---
title: "family Network EDA"
author: "Bijan Ranjbar-Sahraei"
date: "October 25, 2015"
output: html_document
---

A simple R markdown to do exploratory data analysis on Family Networks.

# Import Data
We start by importing the first csv file.

*Please not that I had to add a "no" column to this csv file.*
```{r}
file_name <- "/Users/bian/Dropbox/CATCH_Local/network_data/family_networks_1.csv"
raw_data <- read.csv(file_name,sep=';')
```

Here are columns of *raw_data*
```{r}
colnames(raw_data)
```

Now let's leave out the part of data we don't need and keep the rest in *data*
```{r}
data <- raw_data[,c("network_no","source","target")]
head(data)
```

# Statistics

Let's get size of each network

```{r}
network_size <- matrix(table(data$network_no))
head(network_size)
summary(network_size)
```
So, in average, each network has a size of 19.74 nodes. The smallest network has 9 nodes and the largest one has 26 nodes.
Let's draw a histogram.

```{r}
qplot(data$network_no,binwidth=1)
```

Now it's time to build graphs and compute the degree of each graph, etc.
